---
title: "MXC_MainFigures"
author: "Jessica Gronniger"
date: "2026-04-10"
output: html_document
---

#Load required packages
```{r, warning=FALSE, message=FALSE, results=FALSE}
library(devtools)
library(sjPlot)
library(RColorBrewer)
library(tidyr)
library(dplyr)
library(reshape2)
library("vegan")
library("gridExtra")
library(ggrepel)
library(grid)
library(data.table)
library(plyr)
library(ggpubr)
library(hillR)
library(lmerTest)
library(tidyverse)
library(cowplot)
library(ggplot2)
library(ggeffects) 
library(patchwork)   
library(svglite)
library(phyloseq)
library(ANCOMBC)
```

#color scheme for bar plots (i.e. colors for each isolate)
```{r}

col28 <- c(
  "S9"        = "dodgerblue2",      
  "S33"       = "limegreen",           
  "S8"        = "saddlebrown",       
  "S120"      = "lightblue1",        
  "S14"       = "navy",              
  "S29"       = "mediumorchid1",     
  "S30"       = "darkorchid4",       
  "S21"       = "hotpink",           
  "S11"       = "darkcyan",          
  "S5"        = "paleturquoise1",    
  "S16"       = "skyblue3",          
  "S7"        = "deeppink4",         
  "S32"       = "mediumvioletred",   
  "S6"        = "plum3",             
  "S22"       = "lightpink2",        
  "S491"        = "purple4",          
  "S31"       = "palegreen2",        
  "S27"       = "darkgreen",         
  "S10"       = "mediumaquamarine",         
  "S391"        = "mistyrose2",        
  "S34"       = "forestgreen",       
  "S263"        = "thistle1",      
  "S17"       = "goldenrod3",       
  "S28"       = "lightgoldenrod1",  
  "S2"        = "darkorange3",     
  "S19"       = "salmon",            
  "Unassigned"  = "gray70",            
  "Rare"        = "gray30"             
)


```

#Color scheme for treatments
```{r}

# Assign a unique color for each sample (grouped based on whether it contains "CR", "M9" or "R2A")
color <- c("SM9RTNB" = "#FFC1E3",
           "SM9RTB" = "#CC3399", 
           "SM9VTNB" = "#FF3399",
           "SM9VTB" = "#800040",
           
           "NSM9RTNB" = "#D1B3FF",
           "NSM9RTB" = "#9966CC",
           "NSM9VTNB" = "#4B0082",
           "NSM9VTB" = "#2E003E",
           
          "SR2ARTNB" = "#FFEB3B",
           "SR2ARTB" = "#FBC02D", 
           "SR2AVTNB" = "#FF8C00",
           "SR2AVTB" = "#993D00",
           
           "NSR2ARTNB" = "#FF9999",
           "NSR2ARTB" = "#FF3333",
           "NSR2AVTNB" = "#B22222",
           "NSR2AVTB" = "#660000",
           
           "SCRRTNB" = "#99CCFF",
           "SCRRTB" = "#99FFFF", 
           "SCRVTNB" = "#0066CC",
           "SCRVTB" = "#003366",
           
           "NSCRRTNB" = "#66CC00",
           "NSCRRTB" = "#00FF33",
           "NSCRVTNB" = "#006600",
           "NSCRVTB" = "#003300",
          
          "CR" = "darkgreen",
          "R2A" = "orange",
          "M9" = "purple"
)

```

#############################################################################################

#Figure 2 -  ENS boxplot by treatment and regression


```{r}
#Read in cleaned and organized ASV table
pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable_OGperSample.csv", header=T)


#specify plot labels for faceting

facet_labels <- c("PMXC6" = "Experiment 2", 
                  "PMXC7" = "Experiment 1")


#Linear regression on Environmental Complexity and ENS (hill1)

#Subset data
t12 = pmxc[pmxc$Time == "T12",]

t12_6 = t12[t12$Experiment == "PMXC6",]

t12_7 = t12[t12$Experiment == "PMXC7",]


##Experiment 1 linear regression

fit_linear_7 <- lmer(
  hill1 ~ ComplexityB + (1 | Treatment),
  data = t12_7,
  REML = FALSE
)

pred_linear_7 <- ggpredict(fit_linear_7, terms = "ComplexityB [all]")

r2_linear_7 <- performance::r2_nakagawa(fit_linear_7)$R2_marginal
pval_linear_7 <- anova(fit_linear_7)$`Pr(>F)`[1]

regression7 <- ggplot(t12_7, aes(x = ComplexityB, y = hill1)) +
  geom_point(aes(color = Treatment)) +
  geom_line(
    data = pred_linear_7,
    aes(x = x, y = predicted),
    linewidth = 1
  ) +
  geom_ribbon(
    data = pred_linear_7,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  coord_cartesian(ylim = c(0, 10)) +
  annotate(
    "text",
    x = 0.2, y = 9.5,
    label = paste0(
      "Marginal R² = ", round(r2_linear_7, 3),
      "\n p = ", signif(pval_linear_7, 3)
    ),
    size = 3
  ) +
  facet_wrap(~Experiment, ncol = 1, labeller = as_labeller(facet_labels)) +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 10, margin = margin(t = -30)),
    axis.title.y = element_text(size = 10),
    legend.position = "none",
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) +
  scale_color_manual(values = color) +
  labs(x = "Environmental Complexity", y = "ENS (q=1)")


##Experiment 2 linear regression

fit_linear_6 <- lmer(
  hill1 ~ ComplexityB + (1 | Treatment),
  data = t12_6,
  REML = FALSE
)

pred_linear_6 <- ggpredict(fit_linear_6, terms = "ComplexityB [all]")

r2_linear_6 <- performance::r2_nakagawa(fit_linear_6)$R2_marginal
pval_linear_6 <- anova(fit_linear_6)$`Pr(>F)`[1]

regression6 <- ggplot(t12_6, aes(x = ComplexityB, y = hill1)) +
  geom_point(aes(color = Treatment)) +
  geom_line(
    data = pred_linear_6,
    aes(x = x, y = predicted),
    linewidth = 1
  ) +
  geom_ribbon(
    data = pred_linear_6,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  coord_cartesian(ylim = c(0, 10)) +
  annotate(
    "text",
    x = 0.2, y = 9.5,
    label = paste0(
      "Marginal R² = ", round(r2_linear_6, 3),
      "\n p = ", signif(pval_linear_6, 3)
    ),
    size = 3
  ) +
  facet_wrap(~Experiment, ncol = 1, labeller = as_labeller(facet_labels)) +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 10, margin = margin(t = -30)),
    axis.title.y = element_text(size = 10),
    legend.position = "none",
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) +
  scale_color_manual(values = color) +
  labs(x = "Environmental Complexity", y = "ENS (q=1)")


# Specify desired legend order
t12$Treatment <- factor(t12$Treatment, levels = c("SM9RTNB", "SM9RTB", "SM9VTNB", "SM9VTB", "NSM9RTNB", "NSM9RTB", "NSM9VTNB", "NSM9VTB", "SR2ARTNB", "SR2ARTB", "SR2AVTNB", "SR2AVTB", "NSR2ARTNB", "NSR2ARTB", "NSR2AVTNB", "NSR2AVTB","SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB", "NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))


# Build the plot
legend_plot <- ggplot(t12, aes(x = ComplexityB, y = hill1, color = Treatment)) +
  geom_point() +
  scale_color_manual(values = color) +
  theme_minimal() +
  guides(color = guide_legend(nrow = 24, title.position = "top")) +  # <- title above
  theme(
    legend.position = "right",
    legend.box.margin = margin(0, 0, 0, 0),  # Reduce margin around the entire legend box
    legend.spacing.y = unit(0.5, "pt"),  # Reduce vertical space between legend items
    legend.spacing.x = unit(0.5, "pt"),  # Reduce horizontal space between legend items
    legend.key.size = unit(1.3, "lines"),  # Reduce size of legend keys
    legend.title.align = 0.5  # Center-align title
  )


# Convert to grob
g <- ggplotGrob(legend_plot)

# Extract legend manually
legend_grob <- g$grobs[which(sapply(g$grobs, function(x) x$name) == "guide-box")][[1]]


#Boxplot of diversity metrics by treatment on T12
hill6 = ggplot(t12_6, aes(x=Treatment, y=hill1)) + 
  theme_bw()+
  geom_boxplot(aes(color=Treatment, reorder(Treatment, hill1))) + theme(legend.position="none", axis.text.x = element_text(angle = 90), axis.title.y = element_text(size = 10), axis.title.x = element_text(size = 10)) +ylab("ENS (q=1)") +scale_color_manual(values = color)+
  coord_cartesian(ylim = c(0, 12))+ theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),   # Light gray major grid
    panel.grid.minor = element_line(color = "gray95"),   # Even lighter minor grid
    panel.background = element_blank()) + scale_y_continuous(labels = scales::number_format(accuracy = 1))+ facet_wrap(~Experiment, ncol=1, scales = "free", labeller = as_labeller(facet_labels))

hill7 = ggplot(t12_7, aes(x=Treatment, y=hill1)) + 
  theme_bw()+
  geom_boxplot(aes(color=Treatment, reorder(Treatment, hill1))) +
  coord_cartesian(ylim = c(0, 12))+ theme(legend.position="none", axis.text.x = element_text(angle = 90), axis.title.y = element_text(size = 10), axis.title.x = element_text(size = 10)) +ylab("ENS (q=1)") +scale_color_manual(values = color)+ theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),   # Light gray major grid
    panel.grid.minor = element_line(color = "gray95"),   # Even lighter minor grid
    panel.background = element_blank()) +scale_y_continuous(labels = scales::number_format(accuracy = 1))+ facet_wrap(~Experiment, ncol=1, scales = "free", labeller = as_labeller(facet_labels))


#Arrange plots and save as svg
combined_plot = ggarrange(regression7, regression6,hill7, hill6,  ncol = 2, nrow=2, align="hv", labels = c("A", "B",  "C", "D"), hjust=0.2, vjust=0.6, font.label = list(size = 15,family="serif")) 

combined_with_padding <- ggdraw() +
  draw_plot(combined_plot, x = 0, y = 0, width = 1, height = 1) +
  theme(plot.margin = margin(t = 10, r = 0, b = 0, l = 10, unit = "mm"))

final_plot <- cowplot::plot_grid(
  combined_with_padding,
  legend_grob,
  ncol = 2,
  rel_widths = c(3, 1)
)


svg("~/Desktop/Figure2_Hill1_T12.svg", height=8, width=12)

final_plot

dev.off()
```
#############################################################################################

#Figure 3 - GLM results observed and marginal

```{r}

#S.R2A.RT.NB as reference
pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable_OGperSample.csv", header=T)

t12 = pmxc[pmxc$Time == "T12",]

t12_6 = t12[t12$Experiment == "PMXC6",]

t12_7 = t12[t12$Experiment == "PMXC7",]

#Set reference levels for model
t12_6$Media <- factor(t12_6$Media, levels = c("R2A", "CR"))
t12_6$Temp <- factor(t12_6$Temp, levels = c("RT", "VT"))
t12_6$Shake <- factor(t12_6$Shake, levels = c("S", "NS"))
t12_6$Beads <- factor(t12_6$Beads, levels = c("NB", "B"))

t12_7$Media <- factor(t12_7$Media, levels = c("M9", "CR"))
t12_7$Temp <- factor(t12_7$Temp, levels = c("RT", "VT"))
t12_7$Shake <- factor(t12_7$Shake, levels = c("S", "NS"))
t12_7$Beads <- factor(t12_7$Beads, levels = c("NB", "B"))


#Run GLM on T12 usig Hill1 calculated with 1% abundance cutoff 

#PMXC7 (Experiment 1)

#Run GLM model
glm_pmxc7T12 <- glm(hill1  ~Media*Shake*Temp*Beads, data=t12_7,family = Gamma(link = "log"))

summary(glm_pmxc7T12)

glm_plot7 = plot_model(glm_pmxc7T12,show.values = TRUE, show.p=TRUE, value.offset=0.4,vline="black",transform="exp", type="est")


glm_plot7 = glm_plot7 +theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"))+scale_color_manual(values = c("black","black")) + theme(text = element_text(size = 10)) + ggtitle("M9.CR")+
  coord_flip(clip = "off") + scale_y_log10(limits = c(0.1, 10))

glm_plot_int7 = plot_model(glm_pmxc7T12,show.values = TRUE, show.p=TRUE, value.offset=0.3,vline="black", type="int")


glm_int7 <- glm_plot_int7[[11]]

plot_data7 <- glm_int7$data

#Create a new 'Media' column based on x
plot_data7$Media <- factor(ifelse(plot_data7$x == 1, "M9", "CR"), levels = c("M9", "CR"))

#Create the plot colored by Media
glm_plot_int7 <- ggplot(plot_data7, aes(x = group, y = predicted, color = Media)) +
  geom_point(position = position_dodge(width = 0.3), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, position = position_dodge(width = 0.3)) +
  facet_grid(panel ~ facet) +
  theme_bw() +
  scale_color_manual(values = c("M9" = "purple", "CR" = "darkgreen")) +
  theme(
    axis.line = element_line(colour = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    strip.background = element_rect(fill = "grey90", colour = "black")
  ) + ylab("ENS (q=1)") +xlab("Shake") + ggtitle("Effect plots for M9.CR")


#PMXC6 (Experiment 2)

#Run GLM model
glm_pmxc6T12 <- glm(hill1  ~ Media*Shake*Temp*Beads, data=t12_6,family = Gamma(link = "log"))

summary(glm_pmxc6T12)


glm_plot6 = plot_model(glm_pmxc6T12,show.values = TRUE, show.p=TRUE, value.offset=0.4,vline="black",transform="exp", type="est")


glm_plot6 = glm_plot6 +theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())+scale_color_manual(values = c("black","black")) + theme(text = element_text(size = 10)) + ggtitle("R2A.CR") + coord_flip(clip = "off")+ scale_y_log10(limits = c(0.1, 10))

glm_plot_int6 = plot_model(glm_pmxc6T12,show.values = TRUE, show.p=TRUE, value.offset=0.3,vline="black", type="int")


glm_int6 <- glm_plot_int6[[11]]

plot_data6 <- glm_int6$data

# Step 2: Create a new 'Media' column based on x
plot_data6$Media <- factor(ifelse(plot_data6$x == 1, "R2A", "CR"), levels = c("R2A", "CR"))

# Step 3: Create the plot colored by Media
glm_plot_int6 <- ggplot(plot_data6, aes(x = group, y = predicted, color = Media)) +
  geom_point(position = position_dodge(width = 0.3), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, position = position_dodge(width = 0.3)) +
  facet_grid(panel ~ facet) +
  theme_bw() +
  scale_color_manual(values = c("R2A" = "orange", "CR" = "darkgreen")) +
  theme(
    axis.line = element_line(colour = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    strip.background = element_rect(fill = "grey90", colour = "black")
  )+ ylab("ENS (q=1)")+xlab("Shake") + ggtitle("Effect plots for R2A.CR")


#svglite("~/Desktop/Figure3_GLM_estimates_intPred_T12_min0.01_stacked_v2.svg", height=10, width=12)

layout_matrix <- matrix(c(1,1,1,2, 2, 
                         3,3,3,4,4), 
                       nrow = 2, 
                       byrow = TRUE)


grid.arrange(glm_plot7, glm_plot_int7,glm_plot6, glm_plot_int6, layout_matrix= layout_matrix)


#dev.off()
```

#############################################################################################

#Figure 4 - Final community composition

#Calculate bray curtis with relative abundance
```{r}
pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable.csv", header=T)

#Subset by experiment and set treatment order for barplots

pmxc$Day = as.factor(pmxc$Day)

pmxc6 = pmxc[pmxc$Experiment == "PMXC6",]

pmxc6$Treatment <- factor(pmxc6$Treatment, levels = c("R2A", "CR", "SR2ARTNB", "SR2ARTB", "SR2AVTNB", "SR2AVTB","NSR2ARTNB", "NSR2ARTB", "NSR2AVTNB", "NSR2AVTB", "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))

pmxc7 = pmxc[pmxc$Experiment == "PMXC7",]

pmxc7$Treatment <- factor(pmxc7$Treatment, levels = c("M9", "CR", "SM9RTNB", "SM9RTB", "SM9VTNB", "SM9VTB","NSM9RTNB", "NSM9RTB", "NSM9VTNB", "NSM9VTB", "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))


#Barplot final samples 
t12_6=pmxc6[pmxc6$Time=="T12",]
t12_7=pmxc7[pmxc7$Time=="T12",]

#Melt data into long format for barplot
long6 = reshape2::melt(t12_6, variable.name="Isolate", id.vars = c("Sample", "Time",  "Shake", "Media", "Temp", "Beads", "Rep", "Treatment", "ComplexityB", "Experiment", "ID", "Day", "hill0", "hill1", "hill2"))

long6$Isolate <- as.character(long6$Isolate)

#Melt data into long format for barplot
long7 = reshape2::melt(t12_7, variable.name="Isolate", id.vars = c("Sample", "Time",  "Shake", "Media", "Temp", "Beads", "Rep", "Treatment", "ComplexityB", "Experiment", "ID", "Day", "hill0", "hill1", "hill2"))

long7$Isolate <- as.character(long7$Isolate)

# Define the desired order of ASVs
ordered_asvs6 <- c("S9", "S33", "S8", "S120", "S14", 
                   "S29", "S30", "S21", "S11", "S5", "S16", "S7", "S32", "S6", "S22", "S491", "S31", "S27", "S10", "S391", "S34", "S263", "S17", "S28", "S2", "S19", "Unassigned", "Rare") 

ordered_asvs7 <- c("S9", "S8", "S120", "S14", 
                   "S29", "S30", "S21", "S11", "S5", "S16", "S7", "S32", "S6", "S22", "S491", "S31", "S27", "S10", "S391", "S34", "S263", "S17", "S28", "S2", "S19", "Unassigned", "Rare") 

# Set the factor levels in desired order
long6$Isolate <- factor(long6$Isolate, levels = ordered_asvs6)
long7$Isolate <- factor(long7$Isolate, levels = ordered_asvs7)


comp6 = ggplot(data = long6, aes(fill=Isolate, y=value, x=Rep)) +    
  theme_bw()+        
    geom_bar(position="fill", stat="identity") +
   scale_fill_manual(values = col28) + theme(strip.text = element_text(
    size =10)) +theme(axis.text=element_text(size=15), legend.text = element_text(size=15),legend.key.size = unit(5, 'mm'), axis.text.x = element_text(angle = 90),  axis.title.x=element_blank(), plot.margin = margin(t = 20, r = 20, b = 20, l = 20, unit = "pt"),
    axis.title.y = element_text(size = 15))+ ylab("Relative Abundance") + guides(fill=guide_legend(ncol =1)) + facet_wrap(~Treatment)+ theme(legend.position="none", strip.text = element_text(size = 15))+ ggtitle("R2A.CR")

comp7 = ggplot(data = long7, aes(fill=Isolate, y=value, x=Rep)) +    
  theme_bw()+        
    geom_bar(position="fill", stat="identity") +
   scale_fill_manual(values = col28) + theme(strip.text = element_text(
    size =10)) +theme(axis.text=element_text(size=15), legend.text = element_text(size=15),legend.key.size = unit(5, 'mm'), axis.text.x = element_text(angle = 90),  axis.title.x=element_blank(), plot.margin = margin(t = 20, r = 20, b = 20, l = 20, unit = "pt"), axis.title.y = element_text(size = 15))+ ylab("Relative Abundance") + guides(fill=guide_legend(ncol =1)) + facet_wrap(~Treatment)+ theme(legend.position="none", strip.text = element_text(size = 15)) + ggtitle("M9.CR")


#Extract legend for barplot

comp_legend_plot = ggplot(data = long6, aes(fill=Isolate, y=value, x=Rep)) +    
  theme_bw()+        
    geom_bar(position="fill", stat="identity") +
   scale_fill_manual(values = col28) + theme(strip.text = element_text(
    size =8)) +theme(axis.text=element_text(size=8), legend.text = element_text(size=10), axis.text.x = element_text(angle = 90),  axis.title.x=element_blank(), plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"))+ ylab("Relative Abundance") + guides(fill = guide_legend(ncol = 2)) + facet_wrap(~Treatment)+ theme(
    legend.position = "right",
    legend.box.margin = margin(0, 0, 0, 0),  # Reduce margin around the entire legend box
    legend.spacing.y = unit(0.5, "pt"),  # Reduce vertical space between legend items
    legend.spacing.x = unit(1, "pt"),  # Reduce horizontal space between legend items
    legend.key.size = unit(3, "lines"),  # Reduce size of legend keys
    legend.title.align = 0.5,
    legend.text = element_text(size = 15),        # <- Increase legend text size
    legend.title = element_text(size = 15) 
    )



# Convert to grob
g <- ggplotGrob(comp_legend_plot)

# Extract legend manually
comp_legend <- g$grobs[which(sapply(g$grobs, function(x) x$name) == "guide-box")][[1]]

#svg("~/Desktop/Figure4_T12composition_nmds_v2.svg", width=25, height=20)
ggarrange(nmdsplot7, nmdsplot6, nmds_legend, comp7, comp6, comp_legend, align = "hv")
#dev.off()
```

#############################################################################################

#Figure 5 - ANCOMBC dot plot

```{r}
data6 = read.csv("~/Desktop/PMXC6_ASVtable_rawCounts_notRare_summedASVs_metadata.csv", header=T)

data6 = data6[data6$Time == c("T12"),]

abun6 = data6[,2:28]
metadata6 = data6[,29:37]

rownames(metadata6) = data6$Sample
rownames(abun6) = data6$Sample


tabun6 = t(abun6)

OTU6 = otu_table(as.matrix(tabun6), taxa_are_rows = TRUE)
physeq6 = phyloseq(OTU6)

taxmat6 = matrix(sample(letters, 70, replace = TRUE), nrow = nrow(OTU6), ncol = 7)
rownames(taxmat6) <- rownames(OTU6)
colnames(taxmat6) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
taxmat6

TAX6 = tax_table(taxmat6)

physeq6 = phyloseq(OTU6, TAX6)
physeq6


meta6 = sample_data(metadata6)

meta6$Shake <- factor(meta6$Shake, levels=c("S", "NS"))
meta6$Media <- factor(meta6$Media, levels=c("R2A", "CR"))
meta6$Temp <- factor(meta6$Temp, levels=c("RT", "VT"))
meta6$Beads <- factor(meta6$Beads, levels=c("NB", "B"))
meta6$Treatment <- factor(meta6$Treatment, levels=c("SR2ARTNB", "SR2AVTNB", "SR2AVTB", "SR2ARTB", "SCRVTNB", "SCRVTB", "SCRRTNB", "SCRRTB", "NSR2AVTNB", "NSR2AVTB", "NSR2ARTNB", "NSR2ARTB", "NSCRVTNB", "NSCRVTB", "NSCRRTNB", "NSCRRTB"))


physeq6 = merge_phyloseq(physeq6, meta6)
physeq6


output6 = ancombc2(data = physeq6, assay_name = "counts", tax_level = NULL,
                       fix_formula = "Media * Shake * Temp * Beads",
                       p_adj_method = "fdr", pseudo_sens = TRUE,
                       prv_cut = 0.1,#change as needed, here prvcut set to only include ASV in 10 percent of samples, default is 30% 
                       alpha = 0.05, n_cl = 3, verbose = TRUE)




output6_res <- output6$res
dim(output6_res)

#Ensuring all ASVs are included on y-axis
full_asv_list <- data.frame(taxon = taxa_names(physeq6))

full_res6 <- full_asv_list %>%
  left_join(output6_res, by = "taxon")

full_res6 <- full_res6 %>%
  mutate(across(starts_with("diff_"), ~replace_na(., FALSE))) %>%
  mutate(across(starts_with("lfc_"), ~replace_na(., NA_real_)))


# Step 1: Prepare LFC, SE, significance
lfc_df <- full_res6 %>%
  select(taxon, starts_with("lfc_")) %>%
  pivot_longer(-taxon, names_to = "term", values_to = "lfc") %>%
  mutate(term = str_remove(term, "^lfc_"))

sig_df <- full_res6 %>%
  select(taxon, starts_with("diff_")) %>%
  pivot_longer(-taxon, names_to = "term", values_to = "significant") %>%
  mutate(term = str_remove(term, "^diff_"))

# Step 2: Merge, replace non-significant lfc with NA
plot_df6 <- lfc_df %>%
  left_join(sig_df, by = c("taxon", "term")) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    lfc = ifelse(significant, lfc, NA),
    term = str_replace_all(term, ":", " × ")  # for prettier facet labels
  )

# Step 3: Ensure factor levels to include all taxa and terms on axes
plot_df6$taxon <- factor(plot_df6$taxon, levels = unique(plot_df6$taxon))
plot_df6$term <- factor(plot_df6$term, levels = unique(plot_df6$term))


data7 = read.csv("~/Desktop/PMXC7_ASVtable_rawCounts_notRare_summedASVs_metadata.csv", header=T)

data7 = data7[data7$Time == "T12",]

abun7 = data7[,2:27]
metadata7 = data7[,28:36]

rownames(metadata7) = data7$Sample
rownames(abun7) = data7$Sample


tabun7 = t(abun7)

OTU7 = otu_table(as.matrix(tabun7), taxa_are_rows = TRUE)
physeq7 = phyloseq(OTU7)

taxmat7 = matrix(sample(letters, 70, replace = TRUE), nrow = nrow(OTU7), ncol = 7)
rownames(taxmat7) <- rownames(OTU7)
colnames(taxmat7) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
taxmat7

TAX7 = tax_table(taxmat7)

physeq7 = phyloseq(OTU7, TAX7)
physeq7

meta7 = sample_data(metadata7)

meta7$Shake <- factor(meta7$Shake, levels=c("S", "NS"))
meta7$Media <- factor(meta7$Media, levels=c("M9", "CR"))
meta7$Temp <- factor(meta7$Temp, levels=c("RT", "VT"))
meta7$Beads <- factor(meta7$Beads, levels=c("NB", "B"))
meta7$Treatment <- factor(meta7$Treatment, levels=c("SM9RTNB", "SM9VTNB", "SM9VTB", "SM9RTB", "SCRVTNB", "SCRVTB", "SCRRTNB", "SCRRTB", "NSM9VTNB", "NSM9VTB", "NSM9RTNB", "NSM9RTB", "NSCRVTNB", "NSCRVTB", "NSCRRTNB", "NSCRRTB"))


physeq7 = merge_phyloseq(physeq7, meta7)
physeq7

output7 = ancombc2(data = physeq7, assay_name = "counts", tax_level = NULL,
                       fix_formula = "Media * Shake * Temp * Beads",
                       p_adj_method = "fdr", pseudo_sens = TRUE,
                       prv_cut = 0.1,#change as needed, here prvcut set to only include ASV in 10 percent of samples, default is 30% 
                       alpha = 0.05, n_cl = 3, verbose = TRUE)




output7_res <- output7$res
dim(output7_res)

#Ensuring all ASVs are included on y-axis
full_asv_list <- data.frame(taxon = taxa_names(physeq6))

full_res7 <- full_asv_list %>%
  left_join(output7_res, by = "taxon")

full_res7 <- full_res7 %>%
  mutate(across(starts_with("diff_"), ~replace_na(., FALSE))) %>%
  mutate(across(starts_with("lfc_"), ~replace_na(., NA_real_)))

# Step 1: Prepare LFC, SE, significance
lfc_df <- full_res7 %>%
  select(taxon, starts_with("lfc_")) %>%
  pivot_longer(-taxon, names_to = "term", values_to = "lfc") %>%
  mutate(term = str_remove(term, "^lfc_"))

sig_df <- full_res7 %>%
  select(taxon, starts_with("diff_")) %>%
  pivot_longer(-taxon, names_to = "term", values_to = "significant") %>%
  mutate(term = str_remove(term, "^diff_"))

# Step 2: Merge, replace non-significant lfc with NA
plot_df7 <- lfc_df %>%
  left_join(sig_df, by = c("taxon", "term")) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    lfc = ifelse(significant, lfc, NA),
    term = str_replace_all(term, ":", " × ")  # for prettier facet labels
  )

# Step 3: Ensure factor levels to include all taxa and terms on axes
plot_df7$taxon <- factor(plot_df7$taxon, levels = unique(plot_df7$taxon))
plot_df7$term <- factor(plot_df7$term, levels = unique(plot_df7$term))


# Step 1: Convert to relative abundance
physeq_rel6 <- transform_sample_counts(physeq6, function(x) x / sum(x))

# Step 2: Extract OTU table and ensure taxa are in rows
otu_mat6 <- as(otu_table(physeq_rel6), "matrix")
if (!taxa_are_rows(physeq_rel6)) {
  otu_mat6 <- t(otu_mat6)
}

# Step 3: Compute average relative abundance per taxon
avg_abund_df6 <- rowMeans(otu_mat6, na.rm = TRUE) %>%
  enframe(name = "taxon", value = "mean_relative_abundance") %>%
  arrange(desc(mean_relative_abundance))

# Step 1: Convert to relative abundance
physeq_rel7 <- transform_sample_counts(physeq7, function(x) x / sum(x))

# Step 2: Extract OTU table and ensure taxa are in rows
otu_mat7 <- as(otu_table(physeq_rel7), "matrix")
if (!taxa_are_rows(physeq_rel7)) {
  otu_mat7 <- t(otu_mat7)
}

# Step 3: Compute average relative abundance per taxon
avg_abund_df7 <- rowMeans(otu_mat7, na.rm = TRUE) %>%
  enframe(name = "taxon", value = "mean_relative_abundance") %>%
  arrange(desc(mean_relative_abundance))


plot_df6$Experiment <- "Experiment 2"
plot_df7$Experiment <- "Experiment 1"

plot_df6 <- plot_df6 %>%
  left_join(avg_abund_df6, by = "taxon")

plot_df7 <- plot_df7 %>%
  left_join(avg_abund_df7, by = "taxon")

plot_df6 <- plot_df6[!plot_df6$taxon %in% c("Unassigned", "Unassigne"), ]

plot_df7 <- plot_df7[!plot_df7$taxon %in% c("Unassigned", "Unassigne"), ]


ancombc_67 = rbind(plot_df6,plot_df7)


# Step 1: Define global taxon order (sorted or manual)
ancombc_67 <- ancombc_67 %>%
  arrange(taxon) %>%
  mutate(
    taxon = factor(taxon, levels = rev(c("S9", "S33", "S8", "S120", "S14", 
                   "S29", "S30", "S21", "S11", "S5", "S16", "S7", "S32", "S6", "S22", "S491", "S31", "S27", "S10", "S391", "S34", "S263", "S17", "S28", "S2", "S19", "Unassigne", "Unassigned"))),  # lock y-axis order
    term = factor(term, levels = unique(term))      # optional: control x order
  )
ancombc_67 <- ancombc_67[!ancombc_67$taxon %in% c("Unassigned", "Unassigne"), ]

term_labels <- c(
  "MediaCR" = "Cricket Media (CR)",
  "ShakeNS" = "No Shake (NS)",
  "TempVT" = "Variable Temp (VT)",
  "BeadsB" = "Beads(B)",

  "MediaCR × ShakeNS" = "CR*NS",
  "MediaCR × TempVT" = "CR*VT",
  "ShakeNS × TempVT" = "NS*VT",
  "MediaCR × BeadsB" = "CR*B",
  "ShakeNS × BeadsB" = "NS*B",
  "TempVT × BeadsB" = "VT*B",

  "MediaCR × ShakeNS × TempVT" = "CR*NS*VT",
  "MediaCR × ShakeNS × BeadsB" = "CR*NS*B",
  "MediaCR × TempVT × BeadsB" = "CR*VT*B",
  "ShakeNS × TempVT × BeadsB" = "NS*VT*B",
  "MediaCR × ShakeNS × TempVT × BeadsB" = "CR*NS*VT*B"
)


#Dot plot of ANCOMBC results where color = LFC and size = average relative abundance of each ASV at T12 in their respective experiments


#svglite("~/Desktop/Figure5_ANCOMBC_dotPlot_sizeRelAbun.svg", width=8, height=6)

ancom_plot = ggplot(ancombc_67, aes(x = term, y = taxon)) +
  # Background layer: all points (invisible, just to keep layout)
  geom_point(
    shape = 21, color = NA, fill = NA, stroke = 0, alpha = 0, size = 2,
    na.rm = TRUE
  ) +

  # Foreground layer: only significant points
  geom_point(
    data = ancombc_67 %>% filter(significant),
    aes(fill = lfc, size = mean_relative_abundance),
    shape = 21, color = "black", stroke = 0.2, na.rm = TRUE
  ) +

  scale_fill_gradientn(
    colours = c("blue", "#E0E0E0", "red"),
    values = scales::rescale(c(-2, -0.5, 0, 0.5, 2)),
    limits = c(-10, 10),
    oob = scales::squish,
    name = "Log Fold Change",
    guide = guide_colorbar(
      frame.colour = "black",
      ticks.colour = "black",
      frame.linewidth = 0.3,
      ticks.linewidth = 0.3
    )
  ) +

  scale_size(
    range = c(2, 6),
    breaks = c(0.001, 0.01, 0.1, 0.2, 0.3),
    labels = c("<1%", "1%", "10%", "20%", "30%"),
    name = "Mean Relative Abundance"
  ) +

  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(size = 8)
  ) +
  scale_x_discrete(labels = term_labels) +
  facet_wrap(~Experiment) +
  ylab("") + xlab("")


#dev.off()


```


#Heatmap of relevant chitin genes
```{r}

chi <- read.csv("~/Desktop/PMXC_chitin_hits_summary.csv", header = TRUE)

# Rename strains to ASV
strain_labels <- c(
  "M07D5GC13" = "S16", "M09D5C28" = "S21", "M10D5C16" = "S27", "M10D5GC03" = "S29",
  "M10D5GC09" = "S30", "M10D5GC11" = "S31", "M10D5GC19" = "S32", "M07D5C27" = "S391",
  "M09D6C03" = "S491", "M03D5GC02" = "S5", "M03D5GC15" = "S6", "M03D5GC21" = "S7",
  "M05D5C13" = "S9", "M05D5C33" = "S10", "M05D5GC01" = "S11", "M07D5C33" = "S14",
  "M09D5GC07" = "S22", "M03D5GC05" = "S263", "M03D6GC30" = "S8", "M07D5GC17" = "S17",
  "M03D5C18" = "S2", "M10D6GC05" = "S34", "M07D6GC08" = "S19", "M10D5C22" = "S28",
  "M10D6C16" = "S33", "M07D6C12" = "S120"
)

chi <- chi %>%
  mutate(ASV = recode(Strain, !!!strain_labels))


#Set ASV order to match phylogenetic tree

ordered_asvs <- c("S9", "S33", "S8", "S120", "S14", 
                   "S29", "S30", "S21", "S11", "S11_c2", "S5", "S16", "S7", "S32", "S6", "S22", "S491", "S31", "S27", "S10", "S391", "S34", "S263", "S17", "S28", "S2", "S19") 

chi$ASV <- factor(chi$ASV, levels = rev(ordered_asvs))


#Specify type order
ordered_type <- c("Chitinase","Chitin Deacetylase", "Alpha-N-acetylglucosaminidase", "Beta-N-acetylglucosaminidase", "N-acetylglucosamine-6-phosphate deacetylase") 

chi$Gene <- factor(chi$Gene, levels = ordered_type)

#Specify facet order
chi$Family <- factor(chi$Family, levels = c("GH18", "GH19","GH18|GH19", "CE4","GH5", "GH8", "GH89", "GH20", "EC3.5.1.25"))

chi = chi[chi$Gene != "N-acetylglucosamine-6-phosphate deacetylase",]


chit_plot = ggplot(chi, aes(x = Gene, y = ASV, fill = Count)) +
  geom_tile(color = "black") +  # black outline around each tile
  theme_linedraw() +
  scale_fill_gradient(
    low = "lightgreen", high = "darkgreen",
    na.value = "white",  # white for missing data
    name = "Counts"
  ) +
  scale_y_discrete(labels = strain_labels) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5)  # black outline of plot
  ) +
  labs(fill = "Count") 

#Save chitin plot as svg

#svglite("~/Desktop/PMXC_genome_chitin_heatmap.svg", height=10,width=6)
chit_plot
#dev.off()

```

#Boxplot of relative abundances of two Pedobacter species (S17 and S2) in the M9.CR experiment, in SCRVTB and ~22% in SCRVTNB.
```{r}

# Convert OTU/ASV table to matrix
otu_mat7 <- as(otu_table(physeq_rel7), "matrix")
if (!taxa_are_rows(physeq_rel7)) {
  otu_mat7 <- t(otu_mat7)
}

# Convert to dataframe and add ASV column
otu_df7 <- as.data.frame(otu_mat7)
otu_df7$ASV <- rownames(otu_mat7)

# Extract sample metadata
meta_df7 <- as.data.frame(sample_data(physeq_rel7))

# Combine OTU/ASV table with metadata
# Left join so metadata lines up with samples
otu_meta_df7 <- cbind(meta_df7, t(otu_df7[, -ncol(otu_df7)]))

# Subset to samples with the desired treatments
otu_sub <- subset(otu_meta_df7, Treatment %in% c("SCRVTB", "SCRVTNB"))

# Keep only metadata + the two ASVs of interest
otu_sub <- otu_sub[, c("Treatment", "ASV17", "ASV2")]

```

#Abundance boxplot
```{r}

# Convert to long format (ASV + abundance in one column)
otu_long <- reshape2::melt(otu_sub, id.vars = "Treatment", 
                 variable.name = "ASV", 
                 value.name = "Abundance")

# Make boxplot

#svglite("~/Desktop/PMXC_pedo_plot.svg", height=6, width=6)

ggplot(otu_long, aes(x = Treatment, y = Abundance, fill = ASV)) +
  geom_boxplot() +
  theme_bw() +
  scale_fill_manual(
    values = c("ASV17" = "goldenrod3", 
               "ASV2"  = "darkorange3")
  ) +
  ylab("Relative Abundance")

#dev.off()
```
#Boxplot of abundances for S34, S14 and S29 in CR and M9 at T12
```{r}
# Convert OTU/ASV table to matrix
otu_mat7 <- as(otu_table(physeq_rel7), "matrix")
if (!taxa_are_rows(physeq_rel7)) {
  otu_mat7 <- t(otu_mat7)
}

# Convert to dataframe and add ASV column
otu_df7 <- as.data.frame(otu_mat7)
otu_df7$ASV <- rownames(otu_mat7)

# Extract sample metadata
meta_df7 <- as.data.frame(sample_data(physeq_rel7))

# Combine OTU/ASV table with metadata
# Left join so metadata lines up with samples
otu_meta_df7 <- cbind(meta_df7, t(otu_df7[, -ncol(otu_df7)]))



# Keep only metadata + the two ASVs of interest
otu_sub <- otu_meta_df7[, c("Treatment", "ASV34", "ASV14", "ASV29", "Media")]

# Convert to long format (ASV + abundance in one column)
otu_long <- reshape2::melt(otu_sub, id.vars = c("Treatment", "Media"), 
                 variable.name = "ASV", 
                 value.name = "Abundance")
```

# Make boxplot
```{r}
#svglite("~/Desktop/PMXC_media_S34_S14_S29_plot.svg", height=6, width=6)

ggplot(otu_long, aes(x = ASV, y = Abundance, fill = ASV)) +
  geom_boxplot() +
  theme_bw() +
  scale_fill_manual(
    values = c("ASV34" = "forestgreen", 
               "ASV14"  = "navy",
               "ASV29" = "mediumorchid1")
  ) +
  ylab("Relative Abundance") + facet_wrap(~Media)
 
#dev.off()

```

#Testing PERMANOVA, ANOSIM and SIMPER to compare to ANCOMBC2
#Calculate bray curtis with relative abundance
```{r}
pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable.csv", header=T)

#Subset by experiment and set treatment order for barplots

pmxc$Day = as.factor(pmxc$Day)

pmxc = pmxc[pmxc$Day == 39,]

pmxc6 = pmxc[pmxc$Experiment == "PMXC6",]

pmxc6$Treatment <- factor(pmxc6$Treatment, levels = c("R2A", "CR", "SR2ARTNB", "SR2ARTB", "SR2AVTNB", "SR2AVTB","NSR2ARTNB", "NSR2ARTB", "NSR2AVTNB", "NSR2AVTB", "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))

pmxc7 = pmxc[pmxc$Experiment == "PMXC7",]

pmxc7$Treatment <- factor(pmxc7$Treatment, levels = c("M9", "CR", "SM9RTNB", "SM9RTB", "SM9VTNB", "SM9VTB","NSM9RTNB", "NSM9RTB", "NSM9VTNB", "NSM9VTB", "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))
```

```{r}
#Subset ASV abundances and save as matrix
abund_table6 <- as.matrix(pmxc6[,2:29])
abund_table7 <- as.matrix(pmxc7[,2:29])

#Subset metadata and save as matrix
meta_table6 <- pmxc6[,30:43]
meta_table7 <- pmxc7[,30:43]

#Calculate Bray-Curtis dissimilarity
dist_mat6 = vegdist(abund_table6, method = "bray")
dist_mat7 = vegdist(abund_table7, method = "bray")
```

#PERMANOVA
```{r}
permanova6 <- adonis2(dist_mat6 ~ Treatment,
                     data = meta_table6,
                     permutations = 999)



permanova7 <- adonis2(dist_mat7 ~ Treatment,
                     data = meta_table7,
                     permutations = 999)

disp6 <- betadisper(dist_mat6, meta_table6$Treatment)
anova(disp6)

disp7 <- betadisper(dist_mat7, meta_table7$Treatment)
anova(disp7)

print(permanova6)
print(permanova7)
```
#ANOSIM
```{r}

anosim_res6 <- anosim(dist_mat6, meta_table6$Treatment)
summary(anosim_res6)

anosim_res7 <- anosim(dist_mat7, meta_table7$Treatment)
summary(anosim_res7)

```
#SIMPER
```{r}

simper_res6 <- simper(dist_mat6, meta_table6$Treatment)

summary(simper_res6)

simper_res7 <- simper(dist_mat7, meta_table7$Treatment)

summary(simper_res7)

```


#PCOA plot
```{r}

library(vegan)
library(ggplot2)
library(patchwork)

# PCoA (Experiment 6)
pcoa6 <- cmdscale(dist_mat6, eig = TRUE, k = 2)
pcoa_df6 <- as.data.frame(pcoa6$points)
colnames(pcoa_df6) <- c("PC1", "PC2")
pcoa_df6$Treatment <- meta_table6$Treatment

plot6 <- ggplot(pcoa_df6, aes(PC1, PC2, color = Treatment)) +
  geom_point(size = 2) +
  stat_ellipse() +
  scale_color_manual(values=color)+
  theme_bw() +
  labs(title = "PMXC6", x = "PCoA1", y = "PCoA2")

# PCoA (Experiment 7)
pcoa7 <- cmdscale(dist_mat7, eig = TRUE, k = 2)
pcoa_df7 <- as.data.frame(pcoa7$points)
colnames(pcoa_df7) <- c("PC1", "PC2")
pcoa_df7$Treatment <- meta_table7$Treatment

plot7 <- ggplot(pcoa_df7, aes(PC1, PC2, color = Treatment)) +
  geom_point(size = 2) +
  stat_ellipse() +
  scale_color_manual(values=color)+
  theme_bw() +
  labs(title = "PMXC7", x = "PCoA1", y = "PCoA2")

plot6 / plot7

```

#Dispersion plot
```{r}

disp_df6 <- data.frame(
  distance = disp6$distances,
  Treatment = meta_table6$Treatment
)

disp_df7 <- data.frame(
  distance = disp7$distances,
  Treatment = meta_table7$Treatment
)

dplot6 <- ggplot(disp_df6, aes(Treatment, distance)) +
  geom_boxplot() +
  theme_bw() +
  labs(title = "Experiment 2 Dispersion", y = "Distance to centroid")

dplot7 <- ggplot(disp_df7, aes(Treatment, distance)) +
  geom_boxplot() +
  theme_bw() +
  labs(title = "Experiment 1 Dispersion", y = "Distance to centroid")

dplot7 / dplot6
```


#Summary stats table
```{r}

library(broom)

# PERMANOVA tidy
perm6_tab <- broom::tidy(permanova6) %>%
  mutate(Experiment = "PMXC6")

perm7_tab <- broom::tidy(permanova7) %>%
  mutate(Experiment = "PMXC7")

permanova_table <- bind_rows(perm6_tab, perm7_tab)

# Dispersion results
disp_table <- tibble(
  Experiment = c("PMXC6", "PMXC7"),
  F_value = c(
    anova(disp6)$`F value`[1],
    anova(disp7)$`F value`[1]
  ),
  p_value = c(
    anova(disp6)$`Pr(>F)`[1],
    anova(disp7)$`Pr(>F)`[1]
  )
)

# ANOSIM results
anosim_table <- tibble(
  Experiment = c("PMXC6", "PMXC7"),
  R_stat = c(anosim_res6$statistic, anosim_res7$statistic),
  p_value = c(anosim_res6$signif, anosim_res7$signif)
)

permanova_table
disp_table
anosim_table
```


#between vs within treatment variance
```{r}

d6 <- as.matrix(dist_mat6)
d7 <- as.matrix(dist_mat7)

# PMXC6
same6 <- outer(meta_table6$Treatment, meta_table6$Treatment, "==")
diag(same6) <- NA

within6 <- d6[same6]
between6 <- d6[!same6]

# PMXC7
same7 <- outer(meta_table7$Treatment, meta_table7$Treatment, "==")
diag(same7) <- NA

within7 <- d7[same7]
between7 <- d7[!same7]

mean(within6, na.rm = TRUE)
mean(between6, na.rm = TRUE)

mean(within7, na.rm = TRUE)
mean(between7, na.rm = TRUE)

wilcox.test(within6, between6)
wilcox.test(within7, between7)

library(tidyverse)

df6 <- tibble(
  value = c(within6, between6),
  type = c(rep("Within", length(within6)),
           rep("Between", length(between6)))
)

ggplot(df6, aes(type, value)) +
  geom_boxplot() +
  theme_bw() +
  labs(title = "PMXC6", y = "Bray-Curtis dissimilarity")

```


```{r}

sessionInfo()

```
