---
title: "MXC_Isolate_GrowthCurves"
output: html_document
date: "2026-04-01"
---

#### Load Required Packages ####
```{r}
packages_to_load <- c(
  "ggplot2", "vegan", "lme4", "effects", "tidyverse", "growthcurver", "devtools",
  "plater", "janitor", "reshape", "readr", "ggbiplot", "lubridate"
)

# Load and install required packages
for (i in packages_to_load) { #Installs packages if not yet installed
  if (!require(i, character.only = TRUE)) install.packages(i)
}
```

#Read in growth curve data for isolates grown in R2A, M9TC with glucose, and M9TC without glucose and rename columns based on sample IDs in metadata file
```{r}
#R2A
r2a_data = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/R2A/PMXC_IsolateGC_R2A_data.csv")
r2a_meta = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/R2A/PMXC_IsolateGC_R2A_metadata.csv")

#M9TC no glucose
M9ng_data = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/M9TC_noGlucose/PMXC_IsolateGC_M9NoGlucose_data.csv")
M9ng_meta = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/M9TC_noGlucose/PMXC_IsolateGC_M9NoGlucose_metadata.csv")

#M9TC with glucose
M9wg_data = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/M9TC_withGlucose/PMXC_IsolateGC_M9TCglucose_data.csv")
M9wg_meta = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/M9TC_withGlucose/PMXC_IsolateGC_M9TCglucose_metadata.csv")

#Function to rename columns
rename_with_metadata <- function(data, meta, time_col = "Time") {
  
  meta <- meta %>%
    mutate(SampleName = paste(tube_id, rep, sep = "_"))
  
  # keep time + wells
  data <- data[, colnames(data) %in% c(time_col, meta$wells)]
  
  # rename only non-time columns
  non_time <- colnames(data) != time_col
  
  colnames(data)[non_time] <- meta$SampleName[
    match(colnames(data)[non_time], meta$wells)
  ]
  
  return(data)
}


#Rename data columns
r2a_data  <- rename_with_metadata(r2a_data, r2a_meta)
M9ng_data <- rename_with_metadata(M9ng_data, M9ng_meta)
M9wg_data <- rename_with_metadata(M9wg_data, M9wg_meta)

```

#Subtract average blank value from every OD measurement at each timepoint
```{r}

blank_correct <- function(data, time_col = "Time", blank_pattern = "^blank") {
  
  # make names unique (prevents crash)
  names(data) <- make.unique(names(data))
  
  # find blank columns (handles blank, blank.1, blank.2, etc.)
  blank_cols <- grep(blank_pattern, names(data), ignore.case = TRUE, value = TRUE)
  
  if (length(blank_cols) == 0) {
    stop("No blank columns found.")
  }
  
  # compute row-wise blank mean
  blank_mean <- data %>%
    select(all_of(blank_cols)) %>%
    rowMeans(na.rm = TRUE)
  
  data_corrected <- data %>%
    mutate(across(
      -c(all_of(time_col), all_of(blank_cols)),
      ~ .x - blank_mean
    )) %>%
    select(-all_of(blank_cols)) %>%
    mutate(across(
      -all_of(time_col),
      ~ pmax(.x, 0)
    ))
  
  return(data_corrected)
}

#Subtract blank values from samples
r2a_corrected <- blank_correct(r2a_data)
M9ng_corrected <- blank_correct(M9ng_data)
M9wg_corrected <- blank_correct(M9wg_data)

```

#Convert time column from HH:MM:SS to numerical
```{r}

convert_time_to_hours <- function(df, time_col = "Time") {
  
  if (!is.numeric(df[[time_col]])) {
    df[[time_col]] <- hms(df[[time_col]])
    df[[time_col]] <- as.numeric(df[[time_col]], units = "hours")
  }
  
  return(df)
}

r2a_corrected <- convert_time_to_hours(r2a_corrected)
M9ng_corrected <- convert_time_to_hours(M9ng_corrected)
M9wg_corrected <- convert_time_to_hours(M9wg_corrected)
```

#Plot individual growth curves
```{r}

plot_replicates_by_isolate <- function(df, time_col = "Time") {
  
  df_long <- df %>%
    pivot_longer(
      cols = -all_of(time_col),
      names_to = "Sample",
      values_to = "Value"
    ) %>%
    separate(Sample, into = c("Isolate", "Replicate"), sep = "_", remove = FALSE)
  
  ggplot(df_long, aes(x = .data[[time_col]], y = Value, group = Sample)) +
    geom_line(alpha = 0.7) +
    facet_wrap(~Isolate, scales = "free_y") +
    theme_bw() +
    labs(x = "Time (Hours)", y = "OD600")
}

plot_replicates_by_isolate(r2a_corrected)
plot_replicates_by_isolate(M9ng_corrected)
plot_replicates_by_isolate(M9wg_corrected)
```

#Calculate growth metrics like k(carrying capacity), N0(starting population), r(growthrate), sigma(lag time) using Growthcurver
```{r}

r2a_gr <- SummarizeGrowthByPlate(r2a_corrected,bg_correct = "none", plot_fit = TRUE, 
                                 plot_file = "r2a_gr_plots.pdf")

M9ng_gr <- SummarizeGrowthByPlate(M9ng_corrected,bg_correct = "none", plot_fit = TRUE, 
                                 plot_file = "M9ng_gr_plots.pdf")

M9wg_gr <- SummarizeGrowthByPlate(M9wg_corrected,bg_correct = "none", plot_fit = TRUE, 
                                 plot_file = "M9wg_gr_plots.pdf")
```
#Label each dataset by the type of media used and combine into a single dataset
```{r}

#Function to seperate sample ID into sample name and rep number, and adding a column specifying media used
process_growthcurver <- function(df, media_name) {
  
  df %>%
    separate(sample, into = c("Sample", "Rep"), sep = "_") %>%
    mutate(Media = media_name)
}

r2a_df  <- process_growthcurver(r2a_gr,  "R2A")
M9ng_df <- process_growthcurver(M9ng_gr, "M9ng")
M9wg_df <- process_growthcurver(M9wg_gr, "M9wg")

#combine to get all growth data together
combined_df <- bind_rows(r2a_df, M9ng_df, M9wg_df)

#Fix samples names so they all start with "S"
combined_df <- combined_df %>%
  mutate(
    Sample = case_when(
      str_detect(Sample, "^ASV\\d+$") ~ str_replace(Sample, "^ASV", "S"),
      str_detect(Sample, "^\\d+$")    ~ paste0("S", Sample),
      TRUE ~ Sample
    )
  )

#Remove S33 from M9 media run since it did not grow
combined_df <- combined_df %>%
  filter(!(Sample == "S33" & Media %in% c("M9ng", "M9wg")))

```

#summarise data to get average metrics for each isolate
```{r}

summary_df <- combined_df %>%
  group_by(Sample, Media) %>%
  summarise(
    n_reps = dplyr::n(),
    k_mean      = mean(k, na.rm = TRUE),
    k_sd        = sd(k, na.rm = TRUE),
    N0_mean     = mean(n0, na.rm = TRUE),
    N0_sd       = sd(n0, na.rm = TRUE),
    r_mean      = mean(r, na.rm = TRUE),
    r_sd        = sd(r, na.rm = TRUE),
    sigma_mean  = mean(sigma, na.rm = TRUE),
    sigma_sd    = sd(sigma, na.rm = TRUE),
    .groups = "drop"
  )

```

#Export growth metrics as csv files 
```{r}

write.csv(combined_df, "~/Desktop/MXC_Isolate_Growth_Curves/MXC_Isolate_GrowthMetrics_perMedia.csv")

write.csv(summary_df, "~/Desktop/MXC_Isolate_Growth_Curves/MXC_Isolate_GrowthMetrics_perMedia_summarized.csv")

```

#color scheme for isolates 
```{r}

col28 <- c(
  "S9"        = "dodgerblue2",      
  "S33"       = "limegreen",           
  "S8"        = "saddlebrown",       
  "S120"      = "lightblue1",        
  "S14"       = "navy",              
  "S29"       = "mediumorchid1",     
  "S30"       = "darkorchid4",       
  "S21"       = "hotpink",           
  "S11"       = "darkcyan",          
  "S5"        = "paleturquoise1",    
  "S16"       = "skyblue3",          
  "S7"        = "deeppink4",         
  "S32"       = "mediumvioletred",   
  "S6"        = "plum3",             
  "S22"       = "lightpink2",        
  "S491"        = "purple4",          
  "S31"       = "palegreen2",        
  "S27"       = "darkgreen",         
  "S10"       = "mediumaquamarine",         
  "S391"        = "mistyrose2",        
  "S34"       = "forestgreen",       
  "S263"        = "thistle1",      
  "S17"       = "goldenrod3",       
  "S28"       = "lightgoldenrod1",  
  "S2"        = "darkorange3",     
  "S19"       = "salmon",            
  "Unassigned"  = "gray70",            
  "Rare"        = "gray30"             
)


```


#Boxplot of growth rate for each isolate in M9 (with and without added glucose) and R2A
```{r}

#Set strain order to match phylogeny 

ordered_asvs <- c("S9", "S33", "S8", "S120", "S14", 
                   "S29", "S30", "S21", "S11", "S5", "S16", "S7", "S32", "S6", "S22", "S491", "S31", "S27", "S10", "S391", "S34", "S263", "S17", "S28", "S2", "S19") 

combined_df$Sample <- factor(combined_df$Sample, levels = ordered_asvs)

#Plot growth rate

#pdf("~/Desktop/SF1_GrowthRates.pdf", height=8, width=10)
ggplot(combined_df, aes(x = Sample, y = r)) + 
  theme_bw() +
  geom_boxplot(aes(color = Sample)) +  
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 10),
    axis.text = element_text(color = "black"),  # black tick labels
    axis.line = element_line(colour = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_blank()
  ) +
  ylab("Growth Rate (r)") +
  xlab("Isolate") +
  scale_color_manual(values = col28) +
  facet_wrap(~Media, ncol = 1)

#dev.off()
```
#Comparing the relative influence of casaminos as a primary carbon source compared to casamino + glucose media
```{r}

# subset to only M9
m9_df <- combined_df %>%
  filter(Media %in% c("M9ng", "M9wg"))


#remove outlier with near zero value S17
m9_df <- m9_df %>%
  slice(-136)

ggplot(m9_df, aes(x = Sample, y = r)) + 
  theme_bw() +
  geom_boxplot(aes(color = Sample)) +  
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 10),
    axis.text = element_text(color = "black"),  # black tick labels
    axis.line = element_line(colour = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_blank()
  ) +
  ylab("Growth Rate (r)") +
  xlab("Isolate") +
  scale_color_manual(values = col28) +
  facet_wrap(
  ~Media,
  ncol = 1,
  labeller = as_labeller(c(
    M9ng = "M9 + Casaminos",
    M9wg = "M9 + Casaminos + Glucose"
  ))
)
```
#Make summary table of average growth rates in M9 with and without glucose
```{r}
#removed one rep from S7 and another from S17, values were significantly different from the other two reps, which were consistent

cas_effect = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/MXC_Isolate_GrowthMetrics_perMedia_2outliersRemoved.csv")

#Subset to only M9 media
cas_effect <- cas_effect %>%
  filter(Media %in% c("M9ng", "M9wg"))

#Summarize without outliers
cas_summary <- cas_effect %>%
  group_by(Sample, Media) %>%
  summarise(
    n_reps = dplyr::n(),
    k_mean      = mean(k, na.rm = TRUE),
    k_sd        = sd(k, na.rm = TRUE),
    N0_mean     = mean(n0, na.rm = TRUE),
    N0_sd       = sd(n0, na.rm = TRUE),
    r_mean      = mean(r, na.rm = TRUE),
    r_sd        = sd(r, na.rm = TRUE),
    sigma_mean  = mean(sigma, na.rm = TRUE),
    sigma_sd    = sd(sigma, na.rm = TRUE),
    .groups = "drop"
  )

cas_summary_sub <- cas_summary %>%
  select(Sample, Media, r_mean, r_sd)

#Make summary table for comparison

cas_summary_wide <- cas_summary %>%
  mutate(
    r_summary = sprintf("%.3f ± %.3f", r_mean, r_sd)
  )

cas_summary_wide <- cas_summary_wide %>%
  select(Sample, Media, r_summary) %>%
  pivot_wider(
    names_from = Media,
    values_from = r_summary
  )

write.csv(cas_summary_wide, "~/Desktop/MXC_Isolate_Growth_Curves/M9_casaminoEffect_summary.csv")
```