---
title: "PMXC_SupplementalFigures"
author: "Jessica Gronniger"
date: "2026-04-10"
output: html_document
---
#Load packages
```{r}
library(hillR)
library(ggplot2)
library(patchwork)
library(gridExtra)
library(grid)
library(dplyr)
library(vegan)
library(tidyverse)
library(ggeffects)
library(lme4)
library(lmerTest)
library(performance)
library(svglite)
library(ggpubr)


```

#color scheme for bar plots (i.e. colors for each isolate)
```{r}

col28 <- c(
  "S9"        = "dodgerblue2",      
  "S33"       = "limegreen",           
  "S8"        = "saddlebrown",       
  "S120"      = "lightblue1",        
  "S14"       = "navy",              
  "S29"       = "mediumorchid1",     
  "S30"       = "darkorchid4",       
  "S21"       = "hotpink",           
  "S11"       = "darkcyan",          
  "S5"        = "paleturquoise1",    
  "S16"       = "skyblue3",          
  "S7"        = "deeppink4",         
  "S32"       = "mediumvioletred",   
  "S6"        = "plum3",             
  "S22"       = "lightpink2",        
  "S491"        = "purple4",          
  "S31"       = "palegreen2",        
  "S27"       = "darkgreen",         
  "S10"       = "mediumaquamarine",         
  "S391"        = "mistyrose2",        
  "S34"       = "forestgreen",       
  "S263"        = "thistle1",      
  "S17"       = "goldenrod3",       
  "S28"       = "lightgoldenrod1",  
  "S2"        = "darkorange3",     
  "S19"       = "salmon",            
  "Unassigned"  = "gray70",            
  "Rare"        = "gray30"             
)


```

#Color scheme for treatments
```{r}

# Assign a unique color for each sample (grouped based on whether it contains "CR", "M9" or "R2A")
color <- c("SM9RTNB" = "#FFC1E3",
           "SM9RTB" = "#CC3399", 
           "SM9VTNB" = "#FF3399",
           "SM9VTB" = "#800040",
           
           "NSM9RTNB" = "#D1B3FF",
           "NSM9RTB" = "#9966CC",
           "NSM9VTNB" = "#4B0082",
           "NSM9VTB" = "#2E003E",
           
          "SR2ARTNB" = "#FFEB3B",
           "SR2ARTB" = "#FBC02D", 
           "SR2AVTNB" = "#FF8C00",
           "SR2AVTB" = "#993D00",
           
           "NSR2ARTNB" = "#FF9999",
           "NSR2ARTB" = "#FF3333",
           "NSR2AVTNB" = "#B22222",
           "NSR2AVTB" = "#660000",
           
           "SCRRTNB" = "#99CCFF",
           "SCRRTB" = "#99FFFF", 
           "SCRVTNB" = "#0066CC",
           "SCRVTB" = "#003366",
           
           "NSCRRTNB" = "#66CC00",
           "NSCRRTB" = "#00FF33",
           "NSCRVTNB" = "#006600",
           "NSCRVTB" = "#003300",
          
          "CR" = "darkgreen",
          "R2A" = "orange",
          "M9" = "purple"
)

```

######## Supplemental Figure 1 ##########

#SF1 Boxplot of growth rate for each isolate in M9 and R2A
```{r}

#Read in growth rate data for isolates (See MXC_Isolate_GrowthCurves for raw data and calculations)

gr = read.csv("~/Desktop/MXC_Isolate_Growth_Curves/MXC_Isolate_GrowthMetrics_R2A_M9_SF1.csv", header=T)

#Set strain order to match phylogeny 

ordered_asvs <- c("S9", "S33", "S8", "S120", "S14", 
                   "S29", "S30", "S21", "S11", "S5", "S16", "S7", "S32", "S6", "S22", "S491", "S31", "S27", "S10", "S391", "S34", "S263", "S17", "S28", "S2", "S19") 

gr$Isolate <- factor(gr$Isolate, levels = ordered_asvs)

#Plot growth rate

ggplot(gr, aes(x = Isolate, y = r)) + 
  theme_bw() +
  geom_boxplot(aes(color = Isolate)) +  
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 10),
    axis.text = element_text(color = "black"),  # black tick labels
    axis.line = element_line(colour = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_blank()
  ) +
  ylab("Growth rate (r)") +
  xlab("Isolate") +
  scale_color_manual(values = col28) +
  facet_wrap(
  ~Media,
  ncol = 1,
  labeller = as_labeller(c(
    R2A = "R2A",
    M9 = "M9 + Glucose"
  ))
)


#Save plot
ggsave(
  "~/Desktop/SF1_GrowthRates.svg",
  plot = last_plot(),
  device = svglite::svglite,
  width = 8,
  height = 6
)
```



######## Supplemental Figure 2 ##########

#SF2 Temperature plots for each experiment
```{r}

# Read temperature data file
data <- read.csv("~/Desktop/MXC_Temps_SF2.csv", header = TRUE)

# Convert DateTime column to POSIXct
data$DateTime <- as.POSIXct(data$DateTime, format = "%m/%d/%Y %H:%M:%S")

# Extract just the Date
data$Date <- as.Date(data$DateTime)

#Split by experiment
temp6 = data[data$Experiment=="PMXC6",]
temp7 = data[data$Experiment=="PMXC7",]

# Map each unique Date to a Day number (0 to 39)
unique_dates <- sort(unique(temp6$Date))
day_map <- setNames(0:(length(unique_dates)-1), unique_dates)
temp6$Day <- day_map[as.character(temp6$Date)]

unique_dates <- sort(unique(temp7$Date))
day_map <- setNames(0:(length(unique_dates)-1), unique_dates)
temp7$Day <- day_map[as.character(temp7$Date)]

temp6 <- subset(temp6, Light <= 0)
temp7 <- subset(temp7, Light <= 0)

#Plot
T6 = ggplot(temp6, aes(x = DateTime, y = Temperature, color=Temp)) +
  geom_line() +
  labs(title = "Experiment 2",
       x = "Day",
       y = "Temperature (°C)") +
  theme_bw() +
   scale_x_datetime(
    breaks = seq(min(temp6$DateTime), max(temp6$DateTime), by = "48 hours"),
    labels = function(x) {
      day_labels <- temp6$Day[match(x, temp6$DateTime)]
      return(day_labels)
    }
  ) + ylim(15,30) +scale_color_manual(values = c("RT" = "blue", "VT" = "red"))

T7 = ggplot(temp7, aes(x = DateTime, y = Temperature, color = Temp)) +
  geom_line() +
  labs(title = "Experiment 1",
       x = "Day",
       y = "Temperature (°C)") +
  theme_bw() +
  scale_x_datetime(
    breaks = seq(min(temp7$DateTime), max(temp7$DateTime), by = "48 hours"),
    labels = function(x) {
      day_labels <- temp7$Day[match(x, temp7$DateTime)]
      return(day_labels)
    }
  ) +
  ylim(15, 30) +
  scale_color_manual(values = c("RT" = "blue", "VT" = "red"))

ggarrange(T7, T6, ncol=1)

#Save plot
ggsave(
  "~/Desktop/SF2_Temperatures.svg",
  plot = last_plot(),
  device = svglite::svglite,
  width = 10,
  height = 6
)
```

######## Supplemental Figure 3 ##########

#SF3 Alpha Rarefaction curves for 16s rRNA amplicon data
```{r}

# File paths
pmxc7_file <- "~/Desktop/PMXC7_rareCurve_shannon.csv"   # PMXC7 = Experiment 1
pmxc6_file <- "~/Desktop/PMXC6_rareCurve_shannon.csv"      # PMXC6 = Experiment 2

# Function to reshape QIIME rarefaction output
process_rarefaction <- function(file, experiment_label) {
  df <- read_csv(file, show_col_types = FALSE)
  
  # Rename first column to a standard name
  colnames(df)[1] <- "sample_id"
  
  # Find rarefaction columns (works with either hyphens or periods)
  rare_cols <- grep("^depth[-.][0-9]+_iter[-.][0-9]+$", colnames(df), value = TRUE)
  
  df %>%
    dplyr::select(all_of(c("sample_id", rare_cols))) %>%
    tidyr::pivot_longer(
      cols = all_of(rare_cols),
      names_to = "depth_iter",
      values_to = "value"
    ) %>%
    tidyr::extract(
      depth_iter,
      into = c("depth", "iter"),
      regex = "^depth[-.]([0-9]+)_iter[-.]([0-9]+)$",
      remove = TRUE
    ) %>%
    dplyr::mutate(
      depth = as.numeric(depth),
      iter = as.numeric(iter),
      value = as.numeric(value),
      Experiment = experiment_label
    ) %>%
    dplyr::group_by(sample_id, depth, Experiment) %>%
    dplyr::summarise(value = mean(value, na.rm = TRUE), .groups = "drop")
}

# Process both datasets
pmxc7_df <- process_rarefaction(pmxc7_file, "Experiment 1")
pmxc6_df <- process_rarefaction(pmxc6_file, "Experiment 2")

# Combine both
df_all <- bind_rows(pmxc7_df, pmxc6_df)

# Data frame for vertical lines
vlines <- data.frame(
  Experiment = c("Experiment 1", "Experiment 2"),
  depth = c(17841, 7313)
)

ggplot(df_all, aes(x = depth, y = value, group = sample_id)) +
  geom_line(alpha = 0.35) +
  geom_vline(data = vlines, aes(xintercept = depth),
             linetype = "dashed", color = "red") +
  facet_wrap(~Experiment, ncol = 2, scales = "free_x") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  labs(
    x = "Sampling depth (reads per sample)",
    y = "Shannon diversity"
  )

#Save plot
ggsave(
  "~/Desktop/SF3_rarefaction_curves.svg",
  plot = last_plot(),
  device = svglite::svglite,
  width = 10,
  height = 6
)

```

######## Supplemental Figure 4 ##########

#SF4 Boxplot of Hill numbers at T12
```{r}

pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable.csv", header=T)

pmxc = pmxc[pmxc$Time == "T12",]

pmxc7 = pmxc[pmxc$Experiment == "PMXC7",]
pmxc6 = pmxc[pmxc$Experiment == "PMXC6",]

# Plot
H1_7 = ggplot(pmxc7, aes(x = Treatment, y = hill1)) + 
  theme_bw() +
  geom_boxplot(aes(color = Treatment, reorder(Treatment, hill1))) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),       # <-- remove tick labels
    axis.title.x = element_blank(),      # <-- remove axis title
    axis.title.y = element_text(size = 10)
  ) +
  ylab("ENS (q=1)") +
  scale_color_manual(values = color) +
  coord_cartesian(ylim = c(0, 18)) +
  theme(
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_blank()
  ) +
  scale_y_continuous(labels = scales::number_format(accuracy = 1)) +
  ggtitle("Experiment 1")

H1_6 = ggplot(pmxc6, aes(x = Treatment, y = hill1)) + 
  theme_bw() +
  geom_boxplot(aes(color = Treatment, reorder(Treatment, hill1))) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("ENS (q=1)") +
  scale_color_manual(values = color) +
  coord_cartesian(ylim = c(0, 18)) +
  theme(
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_blank()
  ) +
  scale_y_continuous(labels = scales::number_format(accuracy = 1)) +
  ggtitle("Experiment 2")

H0_7 = ggplot(pmxc7, aes(x = Treatment, y = hill0)) + 
  theme_bw() +
  geom_boxplot(aes(color = Treatment, reorder(Treatment, hill1))) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Richness (q=0)") +
  scale_color_manual(values = color) +
  coord_cartesian(ylim = c(0, 18)) +
  theme(
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_blank()
  ) +
  scale_y_continuous(labels = scales::number_format(accuracy = 1))

H0_6 = ggplot(pmxc6, aes(x = Treatment, y = hill0)) + 
  theme_bw() +
  geom_boxplot(aes(color = Treatment, reorder(Treatment, hill1))) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Richness (q=0)") +
  scale_color_manual(values = color) +
  coord_cartesian(ylim = c(0, 18)) +
  theme(
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_blank()
  ) +
  scale_y_continuous(labels = scales::number_format(accuracy = 1))


H2_7 = ggplot(pmxc7, aes(x=Treatment, y=hill1)) + 
  theme_bw()+
  geom_boxplot(aes(color=Treatment, reorder(Treatment, hill1))) + theme(legend.position="none", axis.text.x = element_text(angle = 90), axis.title.y = element_text(size = 10), axis.title.x = element_text(size = 10)) +ylab("Inverse Simpson (q=2)") +scale_color_manual(values = color)+
  coord_cartesian(ylim = c(0, 18))+ theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),   # Light gray major grid
    panel.grid.minor = element_line(color = "gray95"),   # Even lighter minor grid
    panel.background = element_blank()) + scale_y_continuous(labels = scales::number_format(accuracy = 1))+ xlab("Treatment")

H2_6 = ggplot(pmxc6, aes(x=Treatment, y=hill1)) + 
  theme_bw()+
  geom_boxplot(aes(color=Treatment, reorder(Treatment, hill1))) + theme(legend.position="none", axis.text.x = element_text(angle = 90), axis.title.y = element_text(size = 10), axis.title.x = element_text(size = 10)) +ylab("Inverse Simpson (q=2)") +scale_color_manual(values = color)+
  coord_cartesian(ylim = c(0, 18))+ theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(color = "gray90"),   # Light gray major grid
    panel.grid.minor = element_line(color = "gray95"),   # Even lighter minor grid
    panel.background = element_blank()) + scale_y_continuous(labels = scales::number_format(accuracy = 1)) + xlab("Treatment")


List <- list(H1_7, H1_6,
    H0_7, H0_6,
    H2_7, H2_6)


#Plot
wrap_plots(List,ncol = 2,nrow = 3)

#Save plot
ggsave(
  "~/Desktop/SF4_HillNumbers_T12.svg",
  plot = last_plot(),
  device = svglite::svglite,
  width = 10,
  height = 8
)
```

######## Supplemental Figure 5 ##########

#Read in cleaned and organized ASV table
```{r}

pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable_OGperSample.csv", header=T)

small = read.csv("~/Desktop/PMXC6_7_combined_ASVtable.csv", header=T)

# Get the initial abundances for only day 0
# average across the two seed communities
init_comm <- small %>%
  filter(Day == 0) %>%
  select(Sample:Rare, Experiment) %>% 
  pivot_longer(cols=S10:Rare, names_to = "Species", values_to = "Abundance") %>%  
  mutate(Species = as.factor(Species), Experiment = as.factor(Experiment)) %>%
  dplyr::group_by(Experiment, Species) %>%
  dplyr::summarise(mAbundance = mean(Abundance))



#Experiment 1 (PMXC7) complexity-null-model

PMXC7_init <- init_comm %>%
  filter(Experiment == "PMXC7")

rand_PMXC7_init <- as_tibble(matrix(NA, nrow=100, ncol=1000))
names(rand_PMXC7_init) <- as.character(seq(1:1000))

for (x in 1:1000) {
  rand_PMXC7_init[,x] <- sample(PMXC7_init$Species, 100, replace = TRUE, prob = PMXC7_init$mAbundance)
}

rand_PMXC7_rel_abund <- rand_PMXC7_init %>%
  pivot_longer(cols = 1:1000, names_to = "Run", values_to = "Species") %>%
  arrange(Run) %>%
  dplyr::group_by(Run, Species) %>%
  dplyr::summarise(relAbund = n()/100) %>%
  pivot_wider(names_from = Species, values_from = relAbund, values_fill = 0) %>%
  arrange(as.numeric(Run))

 # calculate hill 1 values
hill1_PMXC7 <- hill_taxa(rand_PMXC7_rel_abund[,-1], q = 1) 

# calculate mean
hill1_mean_PMXC7 <- mean(hill1_PMXC7)

# determine lower and upper 95 percent confidence intervals
# for Hill numbers. Since we genereated 1000 samples, we drop the
# lowest 25 and highest 25 to get our 95 percent CI
hill1_sorted_PMXC7 <- sort(hill1_PMXC7)
hill1_lower_PMXC7 <- hill1_sorted_PMXC7[26]
hill1_higher_PMXC7 <- hill1_sorted_PMXC7[975]



#Experiment 2 (PMXC6) complexity-null-model

PMXC6_init <- init_comm %>%
  filter(Experiment == "PMXC6")

# make an empty data frame that we will fill in with the for loop
rand_PMXC6_init <- as_tibble(matrix(NA, nrow=100, ncol=1000))
names(rand_PMXC6_init) <- as.character(seq(1:1000))

# loop through 1000 iterations
# in each iteration, sample our species list with probability equal to
# the relative abundance of each species
# key is to make sure that mAbundance is relative abundance, thats sums to 1
# it thus is equivalent to a cumulative probability distribution.
# EXTENSION: Jess -- for the other null model approach, you'll want to make sure that your
# relative abundance across the two treatments sums to 1. 
for (x in 1:1000) {
  rand_PMXC6_init[,x] <- sample(PMXC6_init$Species, 100, replace = TRUE, prob = PMXC6_init$mAbundance)
}

# Just played with having different species and double checking that
# their total number of times sampled is approximately equal to 
# their relative abundace. I.e. that I didn't make a mistake. Looks good.
#length(which(rand_PMXC6_init == "S2")) / length(rand_PMXC6_init)


# calculating relative abundance in each run of each species
rand_PMXC6_rel_abund <- rand_PMXC6_init %>%
  pivot_longer(cols = 1:1000, names_to = "Run", values_to = "Species") %>%
  arrange(Run) %>%
  dplyr::group_by(Run, Species) %>%
  dplyr::summarise(relAbund = n()/100) %>%
  pivot_wider(names_from = Species, values_from = relAbund, values_fill = 0) %>%
  arrange(as.numeric(Run))

 # calculate hill 1 values
# Jess, make sure I did this correct/how you did it
hill1_PMXC6 <- hill_taxa(rand_PMXC6_rel_abund[,-1], q = 1) 

# calculate mean diversity
hill1_mean_PMXC6 <- mean(hill1_PMXC6)

# determine lower and upper 95 percent confidence intervals
# for Hill numbers. Since we genereated 1000 samples, we drop the
# lowest 25 and highest 25 to get our 95 percent CI
hill1_sorted_PMXC6 <- sort(hill1_PMXC6)
hill1_lower_PMXC6 <- hill1_sorted_PMXC6[26]
hill1_higher_PMXC6 <- hill1_sorted_PMXC6[975]



#Linear regression on Environmental Complexity and ENS (hill1) now including the null model test of sampling effects plotted with your actual data.

t12 <- pmxc[pmxc$Time == "T12", ]

t12_6 <- t12[t12$Experiment == "PMXC6", ]
t12_7 <- t12[t12$Experiment == "PMXC7", ]

###Experiment 1 (PMXC7)

fit_linear_7 <- lmer(
  hill1 ~ ComplexityB + (1 | Treatment),
  data = t12_7,
  REML = FALSE
)

pred_linear_7 <- ggpredict(fit_linear_7, terms = "ComplexityB [all]")

r2_7 <- performance::r2_nakagawa(fit_linear_7)
pval_7 <- anova(fit_linear_7)$`Pr(>F)`[1]

regression7 <- ggplot(t12_7, aes(x = ComplexityB, y = hill1)) +
  geom_point(aes(color = Treatment)) +
  geom_line(data = pred_linear_7, aes(x = x, y = predicted), linewidth = 1) +
  geom_ribbon(
    data = pred_linear_7,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  coord_cartesian(ylim = c(0, 22)) +
  annotate(
    "text",
    x = 0.75, y = 13,
    label = paste0(
      "R²m = ", round(r2_7$R2_marginal, 3),
      "\nR²c = ", round(r2_7$R2_conditional, 3),
      "\np = ", signif(pval_7, 3)
    ),
    size = 3
  ) +
  facet_wrap(~Experiment, ncol = 1, labeller = as_labeller(facet_labels)) +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 10, margin = margin(t = -30)),
    axis.title.y = element_text(size = 10),
    legend.position = "none",
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) +
  scale_color_manual(values = color) +
  labs(x = "Environmental Complexity", y = "ENS (q=1)") +
  geom_hline(yintercept = hill1_mean_PMXC7, color = "black") +
  geom_hline(yintercept = hill1_lower_PMXC7, color = "black", linetype = "dashed") +
  geom_hline(yintercept = hill1_higher_PMXC7, color = "black", linetype = "dashed")

###Experiment 2 (PMXC6)

fit_linear_6 <- lmer(
  hill1 ~ ComplexityB + (1 | Treatment),
  data = t12_6,
  REML = FALSE
)

pred_linear_6 <- ggpredict(fit_linear_6, terms = "ComplexityB [all]")

r2_6 <- performance::r2_nakagawa(fit_linear_6)
pval_6 <- anova(fit_linear_6)$`Pr(>F)`[1]

regression6 <- ggplot(t12_6, aes(x = ComplexityB, y = hill1)) +
  geom_point(aes(color = Treatment)) +
  geom_line(data = pred_linear_6, aes(x = x, y = predicted), linewidth = 1) +
  geom_ribbon(
    data = pred_linear_6,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  coord_cartesian(ylim = c(0, 22)) +
  annotate(
    "text",
    x = 0.5, y = 13,
    label = paste0(
      "R²m = ", round(r2_6$R2_marginal, 3),
      "\nR²c = ", round(r2_6$R2_conditional, 3),
      "\np = ", signif(pval_6, 3)
    ),
    size = 3
  ) +
  facet_wrap(~Experiment, ncol = 1, labeller = as_labeller(facet_labels)) +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 10, margin = margin(t = -30)),
    axis.title.y = element_text(size = 10),
    legend.position = "none",
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) +
  scale_color_manual(values = color) +
  labs(x = "Environmental Complexity", y = "ENS (q=1)") +
  geom_hline(yintercept = hill1_mean_PMXC6, color = "black") +
  geom_hline(yintercept = hill1_lower_PMXC6, color = "black", linetype = "dashed") +
  geom_hline(yintercept = hill1_higher_PMXC6, color = "black", linetype = "dashed")


# combine plots
regression6 + regression7

#Save plot
ggsave("~/Desktop/SF5_InitialNull_ENSregression.svg",
       regression7 + regression6,
       width = 8, height = 4)
regression6 + regression7
dev.off()
```

######## Supplemental Figure 6 ##########

#SF6  Bray-Curtis changes over time
```{r}
#Load and organize data

#ASV table should have ASVs in columns and samples in rows
pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable_OGperSample.csv", header=T)

pmxc7 = pmxc[pmxc$Experiment == "PMXC7",]
pmxc6 = pmxc[pmxc$Experiment == "PMXC6",]

pmxc6 <- pmxc6[pmxc6$Day %in% c(0, 3, 24, 36, 39), ]

###Experiment 1 (PMXC7)

# Prepare ASV table and metadata
asv.table <- pmxc7[, 2:29]
asv.t <- asv.table[order(row.names(asv.table)), ]

md <- pmxc7[, 30:43]
md <- md[order(row.names(md)), ]
stopifnot(all(row.names(asv.t) == row.names(md)))

expfn <- asv.t
md.expf <- md

big <- data.frame()
microcosm <- unique(md.expf$Treatment)

for (i in microcosm) {
  # Subset by treatment
  md_sub <- md.expf[md.expf$Treatment == i, ]
  asv_sub <- expfn[md.expf$Treatment == i, ]
  
  # Order by Day
  ord <- order(md_sub$Day)
  md_sub <- md_sub[ord, ]
  asv_sub <- asv_sub[ord, ]
  
  # Drop empty ASVs
  asv_sub <- asv_sub[, colSums(asv_sub) > 0]
  
  # Compute Bray-Curtis distance matrix
  dist_mat <- as.matrix(vegdist(asv_sub, method = "bray"))
  
  # Compute average BC per Day (relative to all other samples in same treatment)
  day_vals <- unique(md_sub$Day)
  for (d in day_vals) {
    idx <- which(md_sub$Day == d)
    # mean of distances of samples at this day to all other samples in treatment
    avg_bc <- mean(dist_mat[idx, -idx], na.rm = TRUE)  # exclude self-comparisons
    big <- rbind(big, data.frame(
      Treatment = i,
      Day = d,
      AvgBC = avg_bc
    ))
  }
}

# Convert Treatment to factor if desired
big$Treatment = factor(big$Treatment, levels = c("SM9RTNB", "SM9RTB", "SM9VTNB", "SM9VTB","NSM9RTNB", "NSM9RTB", "NSM9VTNB", "NSM9VTB",  "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))

# Plot
BC7 = ggplot(big, aes(x = Day, y = AvgBC, color = Treatment)) +
  geom_point(size = 3) +
  geom_line(aes(group = Treatment), size = 1) +
  theme_bw() +
  labs(x = "Day", y = "Average Bray-Curtis dissimilarity", title = "Experiment 1") +
  theme(
    legend.position = "right",
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )+ facet_wrap(~Treatment)+theme(legend.position = "none")+scale_color_manual(values = color)


###Experiment 2 (PMXC6)

# Prepare ASV table and metadata
asv.table <- pmxc6[, 2:29]
asv.t <- asv.table[order(row.names(asv.table)), ]

md <- pmxc6[, 30:43]
md <- md[order(row.names(md)), ]
stopifnot(all(row.names(asv.t) == row.names(md)))

expfn <- asv.t
md.expf <- md

big <- data.frame()
microcosm <- unique(md.expf$Treatment)

for (i in microcosm) {
  # Subset by treatment
  md_sub <- md.expf[md.expf$Treatment == i, ]
  asv_sub <- expfn[md.expf$Treatment == i, ]
  
  # Order by Day
  ord <- order(md_sub$Day)
  md_sub <- md_sub[ord, ]
  asv_sub <- asv_sub[ord, ]
  
  # Drop empty ASVs
  asv_sub <- asv_sub[, colSums(asv_sub) > 0]
  
  # Compute Bray-Curtis distance matrix
  dist_mat <- as.matrix(vegdist(asv_sub, method = "bray"))
  
  # Compute average BC per Day (relative to all other samples in same treatment)
  day_vals <- unique(md_sub$Day)
  for (d in day_vals) {
    idx <- which(md_sub$Day == d)
    # mean of distances of samples at this day to all other samples in treatment
    avg_bc <- mean(dist_mat[idx, -idx], na.rm = TRUE)  # exclude self-comparisons
    big <- rbind(big, data.frame(
      Treatment = i,
      Day = d,
      AvgBC = avg_bc
    ))
  }
}

# Convert Treatment to factor and specify order
big$Treatment = factor(big$Treatment, levels = c("SR2ARTNB", "SR2ARTB", "SR2AVTNB", "SR2AVTB","NSR2ARTNB", "NSR2ARTB", "NSR2AVTNB", "NSR2AVTB", "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))

# Plot
BC6 = ggplot(big, aes(x = Day, y = AvgBC, color = Treatment)) +
  geom_point(size = 3) +
  geom_line(aes(group = Treatment), size = 1) +
  theme_bw() +
  labs(x = "Day", y = "Average Bray-Curtis dissimilarity", title = "Experiment 2") +
  theme(
    legend.position = "right",
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )+ facet_wrap(~Treatment)+theme(legend.position = "none")+scale_color_manual(values = color)



#Plot
ggarrange(BC7, BC6)

#Save plot
ggsave(
  "~/Desktop/SF6_BCoverTime.svg",
  plot = last_plot(),
  device = svglite::svglite,
  width = 16,
  height = 8
)

```


######## Supplemental Figure 7 ##########

#SF7 NMDS plot of community composition
#Calculate bray curtis with relative abundance
```{r}
pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable.csv", header=T)

#Subset by experiment and set treatment order for barplots

pmxc$Day = as.factor(pmxc$Day)

pmxc6 = pmxc[pmxc$Experiment == "PMXC6",]

pmxc6$Treatment <- factor(pmxc6$Treatment, levels = c("R2A", "CR", "SR2ARTNB", "SR2ARTB", "SR2AVTNB", "SR2AVTB","NSR2ARTNB", "NSR2ARTB", "NSR2AVTNB", "NSR2AVTB", "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))

pmxc7 = pmxc[pmxc$Experiment == "PMXC7",]

pmxc7$Treatment <- factor(pmxc7$Treatment, levels = c("M9", "CR", "SM9RTNB", "SM9RTB", "SM9VTNB", "SM9VTB","NSM9RTNB", "NSM9RTB", "NSM9VTNB", "NSM9VTB", "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB"))


#Subset ASV abundances and save as matrix
abund_table6 <- as.matrix(pmxc6[,2:29])
abund_table7 <- as.matrix(pmxc7[,2:29])

#Subset metadata and save as matrix
meta_table6 <- pmxc6[,30:43]
meta_table7 <- pmxc7[,30:43]

#Calculate Bray-Curtis dissimilarity
nmds6 = metaMDS(abund_table6, distance = "bray")
nmds7 = metaMDS(abund_table7, distance = "bray")


#Organize NMDS scores in data frame with metadata

data.scores6 = as.data.frame(scores(nmds6)$sites)

#add columns to data frame 
data.scores6$Sample = meta_table6$Sample
data.scores6$Time = meta_table6$Time
data.scores6$Media = meta_table6$Media
data.scores6$Treatment = meta_table6$Treatment
data.scores6$Beads = meta_table6$Beads
data.scores6$Shake = meta_table6$Shake
data.scores6$Temp = meta_table6$Temp
data.scores6$Rep = meta_table6$Rep
data.scores6$Day = meta_table6$Day

head(data.scores6)

data.scores7 = as.data.frame(scores(nmds7)$sites)

#add columns to data frame 
data.scores7$Sample = meta_table7$Sample
data.scores7$Time = meta_table7$Time
data.scores7$Media = meta_table7$Media
data.scores7$Treatment = meta_table7$Treatment
data.scores7$Beads = meta_table7$Beads
data.scores7$Shake = meta_table7$Shake
data.scores7$Temp = meta_table7$Temp
data.scores7$Rep = meta_table7$Rep
data.scores7$Day = meta_table7$Day

head(data.scores7)


#NMDS showing only centroids for OG, T0, T12

og_6 = subset(data.scores6, Time == "OG")
t0_6= subset(data.scores6, Time == "T0")
t12_6= subset(data.scores6, Time == "T12")

scores6=rbind(og_6,t0_6,t12_6)

og_7 = subset(data.scores7, Time == "OG")
t0_7= subset(data.scores7, Time == "T0")
t12_7= subset(data.scores7, Time == "T12")

scores7=rbind(og_7,t0_7,t12_7)

#data.scores = na.omit(data.scores)
centroids6 <- aggregate(cbind(NMDS1, NMDS2) ~ Treatment + Day, 
                       data = scores6, 
                       FUN = mean)
centroidOG_6 = og_6
centroid1_6 = centroids6[centroids6$Day=="3",]
centroid2_6 = centroids6[centroids6$Day=="39",]

#Plot the NMDS with centroids

scores6$Day = as.factor(scores6$Day)

nmdsplot6 = ggplot(scores6, aes(x = NMDS1, y = NMDS2), color = Treatment) +
  geom_point(alpha = 0.2, aes(color = Treatment, shape=Day), size=4) +  # Show individual points
  geom_point(data = centroid1_6, aes(x = NMDS1, y = NMDS2, color = Treatment, shape=Day), size = 8) +
  geom_point(data = centroid2_6, aes(x = NMDS1, y = NMDS2, color = Treatment, shape=Day), size = 8)+
  geom_point(data = centroidOG_6, aes(x = NMDS1, y = NMDS2, color = Treatment, shape=Day), size = 8)+ scale_shape_manual(values=c(8,19,17))+
  theme_minimal() +scale_color_manual(values = color) + theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), legend.position="none", legend.box="vertical", legend.margin=margin(),legend.title = element_text(angle = 90, vjust=1), plot.margin = margin(t = 20, r = 20, b = 20, l = 20, unit = "pt"), axis.text=element_text(size=15), axis.title.y = element_text(size = 15), axis.title.x = element_text(size = 15)) + ggtitle("R2A.CR")+ annotate("text", x = Inf, y = -Inf, label = sprintf("Stress = %.2f", nmds6$stress),
           hjust = 1.1, vjust = -0.5, size = 6)

#data.scores = na.omit(data.scores)
centroids_7 <- aggregate(cbind(NMDS1, NMDS2) ~ Treatment + Day, 
                       data = scores7, 
                       FUN = mean)
centroidOG_7 = og_7
centroid1_7 = centroids_7[centroids_7$Day=="3",]
centroid2_7 = centroids_7[centroids_7$Day=="39",]

#Plot the NMDS with centroids
scores7$Day = as.factor(scores7$Day)

nmdsplot7 = ggplot(scores7, aes(x = NMDS1, y = NMDS2), color = Treatment) +
  geom_point(alpha = 0.2, aes(color = Treatment, shape=Day), size=4) +  # Show individual points
  geom_point(data = centroid1_7, aes(x = NMDS1, y = NMDS2, color = Treatment, shape=Day), size = 8) +
  geom_point(data = centroid2_7, aes(x = NMDS1, y = NMDS2, color = Treatment, shape=Day), size = 8)+
  geom_point(data = centroidOG_7, aes(x = NMDS1, y = NMDS2, color = Treatment, shape=Day), size = 8)+ scale_shape_manual(values=c(8,19,17))+
  theme_minimal() +scale_color_manual(values = color) + theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), legend.position="none", legend.box="vertical", legend.margin=margin(),legend.title = element_text(angle = 90, vjust=1), plot.margin = margin(t = 20, r = 20, b = 20, l = 20, unit = "pt"), axis.text=element_text(size=15), axis.title.y = element_text(size = 15), axis.title.x = element_text(size = 15)) + ggtitle("M9.CR") + annotate("text", x = Inf, y = -Inf, label = sprintf("Stress = %.2f", nmds7$stress),
           hjust = 1.1, vjust = -0.5, size = 6)

#Make nmds plot with both pmxc6 and pmxc7 and extract legend

scores = rbind(scores6, scores7)

scores$Day = factor(scores$Day, levels = c("0", "3", "39"))


scores$Treatment = factor(scores$Treatment, levels = c("SR2ARTNB", "SR2ARTB", "SR2AVTNB", "SR2AVTB","NSR2ARTNB", "NSR2ARTB", "NSR2AVTNB", "NSR2AVTB", "SM9RTNB", "SM9RTB", "SM9VTNB", "SM9VTB","NSM9RTNB", "NSM9RTB", "NSM9VTNB", "NSM9VTB",  "SCRRTNB", "SCRRTB", "SCRVTNB", "SCRVTB","NSCRRTNB", "NSCRRTB", "NSCRVTNB", "NSCRVTB", "CR", "R2A", "M9"))


# Build the plot
nmds_legend_plot <- ggplot(scores, aes(x = NMDS1, y = NMDS2), color = Treatment) +
  geom_point(aes(color = Treatment, shape=Day), size=8) +  # Show individual point
  theme_minimal() +scale_color_manual(values = color)  +  
  theme(
    legend.position = "right",
    legend.box.margin = margin(0, 0, 0, 0),  # Reduce margin around the entire legend box
    legend.spacing.y = unit(0.5, "pt"),  # Reduce vertical space between legend items
    legend.spacing.x = unit(1, "pt"),  # Reduce horizontal space between legend items
    legend.key.size = unit(2.5, "lines"),  # Reduce size of legend keys
    legend.title.align = 0.5,
    legend.text = element_text(size = 15),        # <- Increase legend text size
    legend.title = element_text(size = 15)
  )+ scale_shape_manual(values=c(8,19,17))


# Convert to grob
g <- ggplotGrob(nmds_legend_plot)

# Extract legend manually
nmds_legend <- g$grobs[which(sapply(g$grobs, function(x) x$name) == "guide-box")][[1]]


#Plot
ggarrange(nmdsplot7, nmdsplot6, nmds_legend, align = "hv")

#Save plot
ggsave(
  "~/Desktop/SF7_NMDSplotBC.svg",
  plot = last_plot(),
  device = svglite::svglite,
  width = 16,
  height = 16
)
```


######## Supplemental Figure 8 ##########

#SF8 Original Composition plots
```{r}
pmxc = read.csv("~/Desktop/PMXC6_7_combined_ASVtable.csv", header=T)

#Subset to only OG
pmxc = pmxc[pmxc$Time == "OG",]

#Melt data into long format for barplot
long = reshape2::melt(pmxc, variable.name="Isolate", id.vars = c("Sample", "Time",  "Shake", "Media", "Temp", "Beads", "Rep", "Treatment", "ComplexityB", "Experiment", "ID", "Day", "hill0", "hill1", "hill2"))



# Define the desired order of ASVs
ordered_asvs <- c("S9", "S33", "S8", "S120", "S14", 
                   "S29", "S30", "S21", "S11", "S5", "S16", "S7", "S32", "S6", "S22", "S491", "S31", "S27", "S10", "S391", "S34", "S263", "S17", "S28", "S2", "S19", "Unassigned", "Rare") 



# Set the factor levels in desired order
long$Isolate <- factor(long$Isolate, levels = ordered_asvs)
long$Experiment <- factor(long$Experiment, levels = c("PMXC7", "PMXC6"))

facet_labels <- c("PMXC6" = "Experiment 2", 
                  "PMXC7" = "Experiment 1")

ggplot(data = long, aes(fill=Isolate, y=value, x=Media)) +    
  theme_bw()+        
    geom_bar(position="fill", stat="identity") +
   scale_fill_manual(values = col28) + theme(strip.text = element_text(
    size =10)) +theme(axis.text=element_text(size=15), legend.text = element_text(size=15),legend.key.size = unit(5, 'mm'), axis.text.x = element_text(angle = 90),  axis.title.x=element_blank(), plot.margin = margin(t = 20, r = 20, b = 20, l = 20, unit = "pt"),
    axis.title.y = element_text(size = 15))+ ylab("Relative Abundance") + guides(fill=guide_legend(ncol =1)) + facet_wrap(~Experiment, scales="free", labeller = as_labeller(facet_labels))

#Save plot
ggsave(
  "~/Desktop/SF8_SeedCompositionBarplot.svg",
  plot = last_plot(),
  device = svglite::svglite,
  width = 16,
  height = 10
)

```



